---
title: "Week 01"
author: "Mia Akar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Libraries 

library(ggplot2)

# Random Numbers 

1. Genarate a vector of 100 numbers drawn from a random normal distribution w a mean=10 and sd=5

```{r randomnormal}
nvec <- rnorm(100, 10, 5)
```

```{r normalprob}
# probability distribution 
x<- c(0,5,10,15,25)
y <- pnorm(x, mean=10, sd=5)
qplot(x,y)
```

```{r normalprob}
# density function 
x<- c(0,5,10,15,25)
y <- dnorm(x, mean=10, sd=5)
qplot(x,y)
```

2. Generate a vector of 100 numbers drawn from a random poisson distribution w k=15

```{r}
pvec <- rpois(100,15)
```

# 'sample' vs 'rep' vs 'replicate' vs 'seq'

```{r}
sample(c("A","B","C"), 10, replace = TRUE)
#replace = false means each thing can only be sampled once 
#sample pulls as many things you would like from a vector 
```

```{r}
rep(rnorm(3), 3)
#sampling 3 values from rnorm and we are replicating that to be the exact same three values 
```

```{r}
replicate(3,rnorm(3))
#matrix of 3x3. Each time we sample three values, those values are different 
```

# Element (aka Cell) vs vector vs matrix vs data frame

# list objects 

These are common output from R functions like 'lm()' 'princomp()'
```{r}
list(name= "my list", numbers = rnorm(10), letters=sample(c("A","B","C"),10,replace=T))
#each component gets a little $ that we can use to slice the list 
```

# 'print' vs 'paste'
```{r}

print(pvec)
paste(pvec)
#paste changes the vector into a string and you can combine strings 
```


# Graphing with ggplot 

```{r}
xdat <- data.frame(X=rnorm(100),
                   Y=rnorm(100))
ggplot(aes(x=X,y=Y,data=xdat)) + geom_point()
```


# Data managment 

dplyr, join, pivot 

# Custom functions

```{r}
mean(c(1,2,3,4,5))

Average <- function(x) {
  #this is where the function goes
  return(sum(x)/length(x))
}
```

```{r}
Average(rnorm(100))
```

# Flow control

## if/else

```{r}
if(1==1){}
#if it returns true in the normal brackets, it will do the thing in the curly brackets
```

```{r}
x<- 2

if(x > 5){
  print("yes, it's more than 5")
} else{
  if(x >= 2){
    print("between 2 and 5")
  } else{
    print("2 or less")
  }
}

```

```{r}
ifelse(1==1,"t","f")
#same as above but efficient for short comparisons as opposed to more complex/longer statements 
```

## for(){}

```{r}
for(i in 1:10) {
  print(i)
}
  #i can be called anything 
```

```{r}
mydat<- data.frame(
  x=paste("X",c(1:10)),
  y=paste("Y",c(1:10)))

  for(row in c(1:10)){
  print(mydat[row,1])
    print(mydat[row,2])
}
```

```{r}
#we can nest forloops
for(row in c(1:10)){
  for(col in c(1:2)){
    print(mydat[row,col])
  }
  print(row)  
}
```

## while(){}

# model time

